---
title: "01 What you need to know about dplyr 1.0.0"
author: "LJ"
date: "2022-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here)

ggplot2::theme_set(theme_minimal())


```

## Introduction

In this notebook I will practice the new functions of dplyr using the examples on the site [Towards Data Science: What you need to know about dplyr 1.0.0](https://towardsdatascience.com/what-you-need-to-know-about-the-new-dplyr-1-0-0-7eaaaf6d78ac)

The article, written by [Keith McNulty](https://keith-mcnulty.medium.com/) for medium.com, goes over seven key functions of the dplyr, version 1.0.0

-   `tidyselect` helper functions

-   `relocate()`

-   `summarise()` expansion

-   `across()`

-   `rowwise()`



## The `tidyselect` helper functions

`tidyselect` is an underlying package for others such as tidyr or dplyr itself and it is used to create verbs to interact with the dataframe.

These verbs or helper functions can now be used directly into dplyr as well.

Some examples of these verbs or helper functions are: *all_of()*, *one_of()*, *any_of()*, *starts_with()*, *eval_select()* among others.

For more information about `tidyselect`, please [check here.](https://tidyselect.r-lib.org/articles/tidyselect.html)

We will use a version of the mtcars dataset built as tibble for the examples:

```{r create the mtcars tibble}
## Make a copy of mtcars, make a tibble and view it
mtcars_test <- mtcars %>% 
  rownames_to_column(var = "car") %>% 
  as_tibble()
  

mtcars_test %>% 
  head()


```

As a first example, we will use the *starts_with()* function to select some columns that start with d and the car column

```{r using starts_with() to select columns}
mtcars_test %>% 
  select(car, 
         starts_with("d")) %>% 
  head(3)

```

An example with *any_of()*, which selects the variables of the dataset that are only present in the vector with variables to choose from.

```{r example any_of()}
mtcars_test %>% 
  select(any_of(c("mpg", "cat", "car", "cyl", "house"))) %>% 
  head(3)

```

The order of the columns is given by order of the variables in the vector.

If we need to remove the variables in the previous vector c("mpg", "cat", "car", "cyl", "house") that are present in the dataframe, we could use `-any_of()`

```{r example - any_of() to remove variables from dataframe}
mtcars_test %>% 
  select(-any_of(c("mpg", "cat", "car", "cyl", "house"))) %>% 
  head(3)

```

## `relocate()`

`relocate()` is used to indicate a precise order of the columns in the dataframe.

By default `relocate()` will move the column to the first position from the left.

```{r relocate() default action}
## Place carb on the first column from the left
mtcars_test %>% 
  relocate(carb) %>% 
  head(3)
  

```

`relocate()` takes the arguments `.before`, and `.after` to move columns to specific places before and/or after an indicated variable.

```{r `relocate()` with .before and .after arguments}
## Place the column that start with d, after the car column
mtcars_test %>% 
  relocate(starts_with("d"), .after = car) %>% 
  head(3)

```

## `summarise()` expanded

`summarise()` has now been expanded to return vectors from multiiple inputs. Also it can return now dataframes.

Some examples:

```{r summarising a range}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(mpg_range = range(mpg)) # range gets the min and max from a list of values
  

```

Combinin this with `pivot_wider()` can help create table with a min and max column

```{r using pivot_wider() to better tabulte previous df}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(mpg_range = range(mpg)) %>% 
  mutate(name = rep(c("min", "max"), 
                    length(unique(cyl)))) %>% 
  pivot_wider(names_from = name,
              values_from = mpg_range)

```

The previous example can be coded more economically like this:

```{r same example as before but with less code}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(min = min(mpg), max = max(mpg))
  

```

## `across()` for column wrangling

`across()` was introduced to allow more powerful `summarise` actions together with the `tidyselect()` verbs. `across()` operates across multiple columns and applying dplyr functions. It represents a time-saving solution removing the need to use the `*_at()`, or `*_if()` type of functions of previous dplyr versions.

For example to get the mean of various numerical columns of `mtcars_test`

```{r get means of specific columss using across()}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(across(c("mpg", "hp"), mean))

```

`across()` is a selecting function. We can identify which functions to apply, and then select the columns they will be applied to.  This is a substitute for `mutate_if`.

```{r}
dplyr::storms %>% 
  mutate(across(where(is_character), as_factor)) %>% 
  select(name, status)
```

Also multiple functions can be applied if passed on as a list argument.

```{r using across() with various functions}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(across(c("mpg", "hp"),
                   list(mean = mean, 
                        median = median, 
                        sd = sd)))

```

Using `glue()` syntax to add the name of the column and the function applied to the names fo the new columns.

```{r using glue() syntax for column naming}
mtcars_test %>% 
  group_by(cyl) %>% 
  summarise(across(starts_with("d"), 
                   list(mean = mean,
                        median = median,
                        sd = sd,
                        mode = mode), 
                   .names = "{col}_{fn}_summ"))
```

## `rowwise()`

`rowwise()`is included in `dplyr` to allow operations to be done with values horizontally or in rows.  
For examples, calculation the average number of phones per region in the dataset WorldPhones is calculation that 
can be done horizontally without having to write a formula for it, and thus having those results as a column.

```{r using rowwise to calculate the mean 01}
## Let's create our test tibble from WorldPhones dataset
worldphones_df <- WorldPhones %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "year") %>% 
  as_tibble() %>% 
  clean_names()

worldphones_df %>% 
  head()

```
```{r using rowwise to calculate the mean 02}
## create a new column that will have the mean of all these values
worldphones_df %>% 
  rowwise() %>% 
  mutate(avg = mean(c_across(n_amer:mid_amer))) %>% 
  head()

```

The same result with `rowwise()` can be obtained by adding to it as argument the columns we will need to exclude from the calculation.  In this case, "year".

```{r}
worldphones_df %>% 
  rowwise(year) %>%  ## Adding the column "year" as argument excludes it from the calculation and makes the code less wordy
  mutate(avg = mean(c_across())) %>% 
  head()

```




```{r another rowwise() example with a range of numbers 01}
## Create a tibble with four columns and 10 rows with random numbers
test_df <- tibble(
  a = round(100 * runif(10),0),
  b = a * 2,
  c = round(100* 1 / a, 0),
  d = round(c, 0) + 1
)

test_df

```

```{r another rowwise() example with a range of numbers 02}
test_df %>% 
  rowwise() %>% 
  mutate(sum = sum(c_across(a:d)))

```


## Back to Home

[Home](index.html)



































