---
title: "02 Features dplyr 1.0.0"
author: "LJ"
date: "2022-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here)

ggplot2::theme_set(theme_minimal())

```

## Introduction

In this notebook I will practice the new functions of dplyr using the examples on the site [Tychobra My Favorite dplyr 1.0.0 features](https://www.tychobra.com/posts/2020-05-27-new-dplyr-features/)

The article was written by [Nick Merlino from ](https://www.tychobra.com/about/) for Tychobra, makers of the polished R package for Shiny apps.

These are the functions that will be covered here:

-    `across()`

-    `slice()`

-    `nest_by()`

-   `.before` and `after`

In this practice workbook we'll load the `AmesHousing` package for its corresponding dataset.

```{r download the AmesHousing dataset}
## Load the housing data, clean names, and select six columns
pacman::p_load(AmesHousing)

ames_data <- make_ames() %>% ## this normalizes the Ames Housing dataset a bit. 
  janitor::clean_names() %>% 
  select(sale_price, bsmt_fin_sf_1, first_flr_sf, 
         total_bsmt_sf, neighborhood, gr_liv_area)

ames_data %>% 
  head()

```


## `across()`

`across()` was created to supersede the functions that used to have as suffix
`*_at()`, `*_all()` or `*_if()`.  

The data contains measurements for the basements, rooms, and living area of the houses in square feet.
Let's convert that to yards using a function.

```{r function to convert feet to yards}
ft_to_yrds <- function (x) {
  x / 9
}

```

The old way using mutate would have been:

```{r convert feet to yards using mutate}
ames_data %>% 
  mutate_at(.vars = vars(contains("_sf") | contains("area")), 
            .funs = ft_to_yrds) %>% 
  head()

```

Using `across()`

```{r convert feet to yards using across}
ames_data %>% 
  mutate(across(.cols = c(contains("_sf"), contains("area")),
                .fns= ft_to_yrds)) %>% 
  head()

```

`across()` can be also be used in combination with `where`.

```{r use `across()` with `where`}
## Use the log function in dplyr to transform numeric values
ames_data %>% 
  mutate(across(where(is.numeric), log)) %>% 
  head()


```

Together with `summarise()`, the use of `across()` with `where` can be used to empower more the mutate verb.
`across()` has an argument called .names used to rename columns by adding a prefix

```{r using summarise with acros and where}
ames_data %>% 
  group_by(neighborhood) %>% 
  summarise(across(where(is.numeric),
                   mean,
                   .names = "mean_{col}")) %>% # .names creates new names for the columns with this syntax
  head()
  
```

The following examples shows how `across()` combines a `mutate_if`, and a `mutate_at`
in one single,  _elegant_ line of code.

```{r use `across()` to combine mutate_at and mutate_if}
## Change to double the integer variables that ends with "_sf" only.
## The two conditions are: integer(mutate_if), and has "_Sf" mutate_at)
ames_data %>% 
  mutate(across(where(is_integer) & ends_with("_sf"), 
                      as.double)) %>% 
  head()

```

`across()` can also do a mutate_all if we add everything() as argument: `across(everything(),...`

## `slice()`

 `slice()` and its helpers are now the way to select the top or lower values in a set and
 also to select random samples from the data.  
 
  `slice()`supersedes `top_n()`, `sample_n()`, `sample_frac()` functions due their
  confusing interpretation.
  
  Examples:
  
  Old way to grab the five most expensives houses in the Ames Housing dataset
  
```{r old way with top_n to get 5 highest values}
ames_data %>% 
  top_n(n = 5, 
        wt = sale_price)

```
  
  New way to grab the five most expensives houses with `slice`
  
```{r new way with slice to get 5 highest values}
ames_data %>% 
  slice_max(sale_price, 
            n = 5)

```

Which are the 5 cheapest houses?

```{r new way with slice to get 5 lowest values}
ames_data %>% 
   slice_min(sale_price, 
            n = 5)

```

Let's now select 4 random rows.

```{r select random rows from the dataset}
ames_data %>% 
   slice_sample(n = 4)

```

To get a sample based on a proportion of the set (a percentage), we use the prop argument with
`slice_sample`.

```{r}
## Get a random sample of the 0.2% of the rows.  Result is 5 rows.
ames_data %>% 
  slice_sample(prop = 0.002)
  
```

Additionally, `slice_head()`, and `slice_tails()` are used to show the first or last rows


```{r}
AmesHousing::ames_raw %>% View()
```

## `nest_by()`

`nest_by()` works similar to `group_by()` as to the fact that it adds grouped metadata,
and it brings nested columns into the dplyr analysis.  

It transforms the tibble into a rowwise dataframe.  

In the next example we will use the traditional `group_by()` and `summarise()` to 
calculate average sale price by neighborhood.  This shows how these two work well
to get one single value out of the calculation.

```{r use group_by and summarise to get average sale price by neighborhood}
ames_data %>% 
  group_by(neighborhood) %>% 
  summarise(avg_sale_price = mean(sale_price)) %>% 
  ungroup() %>% 
  slice_head(n = 6)
              

```

For more complicated operations on the grouped rows, such as using linear models,
`nest_by()` stores grouped data as lists in a new column called data.

```{r use `nest_by()` to fit a linear model in ames_data}
## First let's see what happens when we nest a column instead of grouping.
nested_ames <- ames_data %>% 
  nest_by(neighborhood)

nested_ames %>% 
  head(n = 6)

```

`nest_by()` has changed the structure of the table and now we have all the other
columns collected in one single column with the rest of the rows and data as a list.

Let's perform a linear regression model on the data column using the square footage
and sale price to each neighborhood.  The model needs to be vectorized, hence the 
use of `list()`

```{r perform a linear model on the nested column}
nested_ames_model <- nested_ames %>% 
  mutate(lin_model_test = list(lm(sale_price ~ gr_liv_area, data = data)))

nested_ames_model %>% 
  head(n=6)

```

Let's check one of the models.  The Edwards neighborhood.

```{r}
edwards_model <- nested_ames_model %>% 
  filter(neighborhood == "Edwards") %>% 
  pull(lin_model_test)


edwards_model 

```


## Back to Home

[Home](index.html)













































































